---
title: "How to RMarkdown"
author: "Jeremy Metha"
date: "`r Sys.Date()`"
output: html_document
params:
  continent: "Oceania" 
  year: "2007"
---



## Markdown Syntax

### Adding sections

### Spacing

For new paragraphs leave a line between...

Each section you want to space out.

While for new lines, 
it's not enough to just press return.  
You need to make sure you end your lines with at least 2 spaces.  
Or
it
won't
work

### adding emphasis

If you want to make things bold, wrap them up in **two asterisks**.  
While for *italics*, you only need one.  
***Using three will net you both!***

and `back-quotes` will get you code-type text

### lists

ordered lists go like so

1. thing
  - subthing
    1. sub subthing
    2. numbered subthing
2. thing
3. thing

a. thing
b. thing
c. thing

other lists

* blah
* blah
* blah
  * sub-blah
    * sub-sub-blah




### blockquotes

> if you have some important text you'd really like to  stick out.  
you can use blockquotes
>
>> it can also be ***nested***

### URLs
<https://r4ds.had.co.nz/>

#### making links out of URLs
[R for data science](https://r4ds.had.co.nz/) is a ***fantastic*** resource to have on hand





```{r}
library(tidyverse)
library(gapminder)
df <- gapminder
colours <- country_colors
df <- cbind(df, colours)
```
## coding in markdown `r mean(df$gdpPercap)`
The countries and populations included in this analysis are:

> `r unique(df$country)`

## plots and figures

```{r, results='asis'}

for (i in unique(df$continent)){
  subset <- df %>% 
    filter(continent == i)
  plot <- ggplot(data = subset, aes(x = gdpPercap, y = lifeExp, size = pop))+
    geom_point()+
    ggtitle(paste("countries in", i, sep = " "))
  print(plot)
  
  cat("\n\n")
    }

ggplot(data = df, aes(x = year, y = pop, colour = continent, group = country))+
  geom_line()
  

```

## tables


```{r}
summaries <- df %>% 
  filter(year == params$year) %>% 
  group_by(continent, year) %>% 
  summarise(LifeExpectancy = mean(lifeExp))

knitr::kable(summaries)

```

## model outputs

```{r}

library(broom)

model <- lm(data = df, lifeExp ~ gdpPercap)

tidy_model <- tidy(summary(model))

tidy_model %>%
  knitr::kable()

```

