---
title: "Gapminder Report"
author: "Jeremy Metha"
date: "`r Sys.Date()`"
output: html_document
params:
  year: "2007"
  continent: "Asia"
  
---


```{r, include = FALSE}
library(tidyverse)
library(broom) # for tidying up model outputs to dataframes
library(gapminder) # the gapminder data we'll be working on
df <- gapminder
colours <- country_colors
df <- cbind(df, colours)

```

## Gapminder dataset

This data contains the following columns from the [gapminder project](https://gapminder.org/data/) 


* country
* continent
  * data comes from these continents `r unique(df$continent)`.
* year
  *  contains the years `r unique(df$year)`.
* lifeExp
* pop
* gdpPercap

## Population growth over time

The two highest lines come from China and India, with populations of `r df %>% filter(year == 2007 & country == "China") %>% select(pop)` and `r df %>% filter(year == 2007 & country == "India") %>% select(pop)` in 2007 respectively.

```{r echo=FALSE}
ggplot(df, aes(x = year, y = pop, group = country, colour = continent))+
  geom_line()+
  scale_y_log10() + theme_bw()

```

## A table of population growth over time on a continent by continent basis

```{r echo = FALSE, message=FALSE}

pop_summaries <- df %>% 
  group_by(continent, year) %>% 
  summarise(meanPopulation = mean(pop)) %>% 
  arrange(meanPopulation)

knitr::kable(pop_summaries)

```

## Linear regression investigating the relationship between population growth and time 
```{r echo=FALSE}

  model <- lm(data = df, pop ~ year)

  tidy_model <- tidy(model)
  knitr::kable(tidy_model)
```




```{r, include = FALSE, echo = FALSE, message=FALSE, results ='asis'}
# you can control whether or not to include this chunk by toggling the include = statement between true and false
# results = 'asis' allkows us to use the cat() function to create sections within a code chunk

cat("## Life expectancy and GDP per captia relationships: 2007 all continents")
data_2007 <- df %>% 
  filter(year == 2007)

cat("### Graph")
ggplot(data_2007, aes(x = gdpPercap, y = lifeExp, size = pop, colour = continent))+
  geom_point()

cat("\n\n") # adds two new lines 

cat("### Table")

life_summaries <- data_2007 %>% 
  group_by(continent) %>% 
  summarise(LifeExpectancy = mean(lifeExp))

knitr::kable(life_summaries)

cat("\n\n")

cat("### Regression")

model <- lm(data = data_2007, lifeExp~ gdpPercap)
tidy_model <- tidy(model)
knitr::kable(tidy_model)

```

## Life expectancy and GDP per captia relationships: Each year for all continents with a `for loop`

```{r, , include = FALSE, echo = FALSE, message=FALSE, results ='asis'}

# Within this framework if we want to generate graphs for each year we can do so in a loop!
for (i in unique(df$year)) {
  subset <- df %>% 
  filter(year == i)

cat("##", i)

cat("\n\n")

cat("### Graph of", i)
plot <- ggplot(subset, aes(x = gdpPercap, y = lifeExp, size = pop, colour = continent))+
  geom_point()
print(plot)

cat("\n\n") # adds two new lines 

cat("### Table of",i)

life_summaries <- subset %>% 
  group_by(continent) %>% 
  summarise(LifeExpectancy = mean(lifeExp))

print(knitr::kable(life_summaries))
cat("\n\n")

cat("### Regression of", i)

model <- lm(data = subset, lifeExp~ gdpPercap)
tidy_model <- tidy(model)
print(knitr::kable(tidy_model))


cat("\n\n\n\n")
}
```

## Life expectancy and GDP per captia relationships: selecting year and continent with `params`
```{r, echo = FALSE, message=FALSE, results ='asis', out.width="100%", out.height="100%"}

  subset <- df %>% 
  filter(year == as.numeric(params$year)) %>% 
  filter(continent == params$continent)

cat("This data comes from ", params$continent, " in the year ", params$year, ".", sep = "")

cat("\n\n")

cat("### Graph")
plot <- ggplot(subset, aes(x = gdpPercap, y = lifeExp, colour = country, size = pop))+
  geom_point(alpha = 0.8) +
  scale_colour_manual(values = subset$colours)+
  theme_bw()+
  guides(size = guide_legend(position = "inside"), 
         country = guide_legend(position = "right")) +
  theme(legend.position.inside = c(0.9, 0.1))
plot

cat("\n\n") # adds two new lines 

cat("### Table")

life_summaries <- subset %>% 
  group_by(country) %>% 
  summarise(LifeExpectancy = mean(lifeExp), Population = mean(pop))

print(knitr::kable(life_summaries))
cat("\n\n")

cat("### Regression")

model <- lm(data = subset, lifeExp~ gdpPercap)
tidy_model <- tidy(model)
print(knitr::kable(tidy_model))

cat("\n\n\n\n")
```
